Capon Beamforming Applied on Moving Objects - An Analysis of Local Shift-Invariance

### 2013-10-25 HRUIofMovingObjects/jrnl_article.tex
Finished 1st draft of introduction in Capon on moving object paper

\begin{document}
%
% paper title
% can use linebreaks \\ within to get better formatting as desired
\title{Capon Beamforming Applied on Moving Objects - An Analysis of Local Shift Invariance}
%
%
% author names and IEEE memberships
% note positions of commas and nonbreaking spaces ( ~ ) LaTeX will not break
% a structure at a ~ so this keeps an author's name from being broken across
% two lines.
% use \thanks{} to gain access to the first footnote area
% a separate \thanks must be used for each paragraph as LaTeX2e's \thanks
% was not built to handle multiple paragraphs
%

@@

\IEEEPARstart{W}{ith} the introduction of a fast implementation of the Capon adaptive beamformer, a stream of images can now be processed at interactive frame rates \cite{Asen}. A natural question then is how a moving object appear after Capon beamforming has been applied. Earlier work on Capon beamforming for medical ultrasound imaging has investigated the method's capabilities in single frames only, and temporal behaviour and effects has therefore been neglected.

An important property of an ultrasound imaging system is its local shift invariance. Local shift invariance means that an object’s appearance should be preserved when the object, or the ultrasound probe, is moved slightly. Local shift invariance is what makes techniques like doppler and flow imaging and speckle tracking possible \comment{(need citation)} .

Capon beamforming has the ability to provide increased lateral resolution and contrast in an ultrasound image \cite{Synnevag2009, Chen2011}. However, it is also sensitive to errors like perturbations of the array element, correlated signals and errors in the assumed steering vector.

\comment{Add intro to other work.}

In this paper we analyze the shift invariance property of the Capon beamformer by studying simulated and in vivo images of moving objects. We show that when imaging moving point scatterers using the Capon beamformer with the same beam spacing as used for delay-and-sum, artifacts arise in the image, manifested as oscillating point scatteres. \comment{(TODO: Process csound data to see if we need over sampling on in vivo data, How often do we have bright point scatterers in vivo? Might improve contrast further)}.

@@

\subsection{Capon Beamforming}
The Capon beamformer produce a weight vector, like the one in (\ref{eq:das}), based on the impinging signal and an optimization criteria. The minimization problem is as follow:
\begin{align}
&\min_{\vec{w}} E\{|z[n]|^2\} \rightarrow \min_{\vec{w}} \vec{w}^H \mat{\hat{R}} \vec{w} \label{eq:capon_optimization_criteria} \\
&\text{subjected to } \vec{w}^H\vec{a} = 1,
\end{align}
where $\mat{\hat{R}}$ is the sample covariance matrix. Hence, the resulting weight vector will minimizes the output power while maintaining unit gain along the steering direction $\vec{a}$. %The covariance matrix is usually unknown and has to be estimated from the element data. The covariance matrix in (\ref{eq:capon_optimization_criteria}) is therefore substituted with the sample covariance matrix.

The sample covariance matrix can be estimated in the following way \cite{Synnevag2009}:
\begin{align}
\mat{\breve{R}}[n] = \frac{1}{N_LN_K}\sum_{n'=n-K}^{n+K} \sum_{l=0}^{N_L-1} \vec{x}_l[n']\vec{x}_l[n']^H,
\end{align}
where $K = M-L+1$ typical is proportional to the pulse length, $N_K = 2K + 1$, $N_L = M-L+1$, and $\vec{x}_l$ is the $l\text{th}$ subarray $[x_l[n], \dotso x_{l+L}[n]]$ of length $L$. Finally $\mat{\hat{R}}$ is loaded with a diagonal factor for numerical stability and increased robustness, 
\begin{align}
\epsilon &= d*\text{trace}\{\mat{\breve{R}}\}/L\\
\mat{\hat{R}} &= \mat{\breve{R}} + \epsilon\mat{I}.
\end{align}

The solution to the minimization problem in (\ref{eq:capon_optimization_criteria}) is
\begin{align}
\vec{w}[n] = \frac{\mat{\hat{R}}[n]^{-1}\vec{a}}{\vec{a}^H\mat{\hat{R}}[n]^{-1}\vec{a}} \in \mathbb{C}^L,
\end{align}.

Note that one matrix has to be constructed and inverted for each data vector $\vec{x}$ received by the system. This is indeed computational demanding, and has for long limited the methods accessibility. Following the innovation in GPU computing, this has now changed.

@@

The lateral sampling of an ultrasound image is given by the aperture size on transmit ($D_{tx}$) and receive ($D_{rx}$) combined and the center frequency ($f_c = c/\lambda$), where $c$ is the speed of sound in the media and $\lambda$ is the wavelength. The required beam spacing is then:
\begin{align}
\Delta < \frac{\lambda}{D_{tx} + D_{rx}}.
\end{align}
If this requirement is not met, aliasing effects will appear in the image and the system will not be lateral shift invariant.

@@

>>>>

\begin{document}
%
% paper title
% can use linebreaks \\ within to get better formatting as desired
\title{Capon Beamforming Applied on Moving Objects - An Analysis of Local Shift-Invariance}
%
%
% author names and IEEE memberships
% note positions of commas and nonbreaking spaces ( ~ ) LaTeX will not break
% a structure at a ~ so this keeps an author's name from being broken across
% two lines.
% use \thanks{} to gain access to the first footnote area
% a separate \thanks must be used for each paragraph as LaTeX2e's \thanks
% was not built to handle multiple paragraphs
%

@@

\IEEEPARstart{W}{ith} the introduction of a fast implementation of the Capon adaptive beamformer, a stream of images can now be processed at interactive frame rates \cite{Asen}. It is therefore an interesting question how a moving object appears after Capon beamforming has been applied. Earlier work on Capon beamforming for medical ultrasound imaging has investigated the method's capabilities in single frames only, and temporal behaviour and effects has therefore been neglected.

An important property of an ultrasound imaging system is its local shift-invariance. Local shift-invariance means that an object’s appearance should be preserved when the object, or the ultrasound probe, is moved slightly. This local movement invariance is what makes ultrasound imaging feasible in general, however it is specially important for quantitative techniques that correlate time-varying features, like pulsed doppler and flow imaging, and speckle tracking of tissue and blood. In this paper we analyze the shift-invariance property of the Capon beamformer by studying the beamformer output when applied on simulated and \textit{in vitro} images of moving objects.

Capon beamforming has been shown to provide increased lateral resolution and contrast in ultrasound images \cite{Synnevag2007, Synnevag2009, Chen2011}. However, it is also known for its suboptimal performance if the sample size available for statistical estimation is limited, if correlated signals are present, and if the assumed steering vector does not match the signal propagation vector\cite{Mestre2006, Widrow1982, Wax1996, Wax1996a}. This has led to the development of several robust versions of the Capon beamformer using either spatial smoothing \cite{Shan1985}, diagonal loading \cite{JianLi2003} or steering vector uncertainty sets \cite{Lorenz2005, Rubsamen2013}. The first two, together with temporal smoothing, have been applied in earlier work on Capon beamforming for medical ultrasound imaging \cite{Synnevag2009}, and both trades resolution for robustness. The latter has not been applied, even though it is well known that the Capon beamformer exhibits severe signal-nulling if the steering vector does not match the signal steering vector exactly. This self-nulling actually contribute a lot to the method's super resolution capability in passive-narrowband applications.

In \cite{Cox1973}, Cox derives formulas for the resolution of the Capon beamformer (called the $\vec{k}_3$ processor), and presents in great detail how the amount of mismatch between the selected steering vector and the signal propagation vector impacts the beamformer output in simple situations. Cox states the following important fact about the Capon beamformer (the $\vec{k}_3$ processor): 
\begin{quote}
"\textit{... a $\vec{k}_3$ processor requires more closely spaced beams than a ... $\vec{k}_1$-processor in order to avoid serious signal suppression effects being introduced on signals arriving from directions between the beams}", 
\end{quote}
where $\vec{k}_1$ is the delay-and-sum (DAS) beamformer. In Fig.\,\ref{fig:das_capon_beams} we have plotted the steered response for DAS and the Capon beamformer using (26) and (32) in \cite{Cox1973}. A 64-element linear array with $\lambda/2$ element spacing, where $\lambda$ is the wavelength, is scanned over a sector equal to three times the system resolution. Impinging on the array is a plane wave with a propagation vector orthogonal to the array surface normal with no spatially white noise present. Hence, the interference-noise matrix $\sigma_0^2\mat{Q}$ is the identity. We have also added the effect of typical sub-array averaging and diagonal loading \cite{Asen} to the Capon beamformer output. The figure shows that when the DAS beamformer is used, only a 3.92 dB scallop loss \cite{Harris1978} should be observed if the source of the plane wave starts moving parallel to the array. However if the Capon beamformer is used with the beam density sufficient for DAS the scallop loss could be as large as $31$ dB.

\begin{figure}[!t]
\centerline{
\includegraphics[width=\linewidth]{\img steerd_respons_capon_das_resolution.eps}
}
\caption{...}
\label{fig:das_capon_beams}
\end{figure}

In previous literature on Capon beamforming for medical ultrasound imaging in single frames this effect has been passed over in silence by either oversampling on transmit, or carefully position point scatterers on transmit-receive beam pairs, or ignorance. In this paper we show that when imaging a moving point scatterer with ultrasound using the Capon beamformer with the same beam spacing used for DAS the same scallop loss is observed, manifested as oscillating point scatteres. The question is then, what can we do in order to improve on the situation.

@@

\subsection{Capon Beamforming}
The Capon beamformer produce a weight vector, like the one in (\ref{eq:das}), based on the impinging signal and an optimization criteria. The minimization problem is as follow:
\begin{align}
&\min_{\vec{w}} E\{|z[n]|^2\} \rightarrow \min_{\vec{w}} \vec{w}^H \mat{\hat{R}} \vec{w} \label{eq:capon_optimization_criteria} \\
&\text{subjected to } \vec{w}^H\vec{a} = 1,
\end{align}
where $\mat{\hat{R}}$ is a sample covariance matrix. Hence, the resulting weight vector will minimizes the output power while maintaining unit gain in the steering direction $\vec{a}$. %The covariance matrix is usually unknown and has to be estimated from the element data. The covariance matrix in (\ref{eq:capon_optimization_criteria}) is therefore substituted with the sample covariance matrix.

The sample covariance matrix can be estimated in the following way \cite{Synnevag2009}:
\begin{align}
\mat{\breve{R}}[n] = \frac{1}{N_LN_K}\sum_{n'=n-K}^{n+K} \sum_{l=0}^{N_L-1} \vec{x}_l[n']\vec{x}_l[n']^H,
\end{align}
where $K = M-L+1$ typical is proportional to the pulse length, $N_K = 2K + 1$, $N_L = M-L+1$, and $\vec{x}_l$ is the $l\text{th}$ subarray $[x_l[n], \dotso x_{l+L}[n]]$ of length $L$. Finally the matrix is loaded with a diagonal factor for numerical stability and increased robustness, 
\begin{align}
\epsilon &= d*\text{trace}\{\mat{\breve{R}}\}/L\\
\mat{\hat{R}} &= \mat{\breve{R}} + \epsilon\mat{I}.
\end{align}

The solution to the minimization problem in (\ref{eq:capon_optimization_criteria}) is
\begin{align}
\vec{w}[n] = \frac{\mat{\hat{R}}[n]^{-1}\vec{a}}{\vec{a}^H\mat{\hat{R}}[n]^{-1}\vec{a}} \in \mathbb{C}^L.
\end{align}

Note that one matrix has to be constructed and inverted for each data vector $\vec{x}[n]$ received by the system. This is indeed computational demanding, and has for long limited the methods accessibility. Following the innovation in GPU computing, and by applying a novel method in order to reduce the size of the matrix inversion problem \cite{Nilsen2009}, this has now changed \cite{Asen}.

@@

The lateral sampling of an ultrasound image is given by the aperture size on transmit ($D_{tx}$) and receive ($D_{rx}$) combined \cite{Hergum2007} and the center frequency ($f_c = c/\lambda$), where $c$ is the speed of sound in the media and $\lambda$ is the wavelength. The required Nyquist beam spacing is then:
\begin{align}
\Delta < \frac{\lambda}{D_{tx} + D_{rx}}.
\end{align}
If this requirement is not fulfilled, aliasing effects will appear in the image and the system will not be lateral shift invariant.

@@

\subsection{Impact on Real Time Performance}
Discuss, and show with benchmarks how the different methods for making Capon beamforming shift-invariant impact the total execution time.
